Cofinality and the Delta System Lemma

### 2020-12-26 Delta_System_Lemma/document/root.tex
better document sub/super-scripts
introduction text to DSL

@@

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

@@

>>>>

\usepackage{relsize}
\DeclareRobustCommand{\isactrlbsub}{\emph\bgroup\math{}\sb\bgroup\mbox\bgroup\isaspacing\itshape\smaller}
\DeclareRobustCommand{\isactrlesub}{\egroup\egroup\endmath\egroup}
\DeclareRobustCommand{\isactrlbsup}{\emph\bgroup\math{}\sp\bgroup\mbox\bgroup\isaspacing\itshape\smaller}
\DeclareRobustCommand{\isactrlesup}{\egroup\egroup\endmath\egroup}

@@

% urls in roman style, theory text in math-similar italics
\urlstyle{tt}
\isabellestyle{it}

@@

\section{Introduction}

The session we present gathers very basic results built on the set
theory formalization of Isabelle/ZF
\cite{DBLP:journals/jar/PaulsonG96}. In a sense, some of the material
formalized here corresponds to a natural continuation of that
work. This is even clearer after perusing Section~\ref{sec:zf-lib},
where notions like cardinal exponentiation are first defined, together
with various lemmas that do not depend on the Axiom of Choice ($\AC$);
the same holds for the basic theory of cofinality of ordinals, which
is developed in Section~\ref{sec:cofinality}. In
Section~\ref{sec:cardinal-lib}, (un)countability is defined
and several results proved, now using $\AC$ freely; the latter is also
needed to prove König's Theorem on cofinality of cardinal
exponentiation. Finally, the simplest infinitary version of the Delta
System Lemma (DSL, also known as the ``Sunflower Lemma'') due to  \v{S}anin
is proved in Section~\ref{sec:dsl}.

A greater part of this development was motivated by a joint project on
the formalization of the ctm approach to
forcing~\cite{2020arXiv200109715G} by Gunther, Pagano, Steinberg, and
the author. Indeed, most of the results presented here are required
for the development of forcing; in particular, DSL is used in the
proof that Cohen posets satisfy \emph{countable chain condition}, which
in turn is used to characterize cardinals in generic extensions
\cite[Chap.~IV]{kunen2011set}. As it turns out, the material as
formalized presently will not be part of the forcing formalization,
since for that goal we need relativized versions of both the
concepts and the proofs.
